The Castro AMR Simulation Code: Current and Future Developments

### 2020-02-08 astronum_2019_Castro/astronum-castro-2019.tex
update table
update table caption

The SDC algorithm used in \castro\ follows the ideas of
\cite{dutt:2000,minion:2003}, and uses low order explicit advection
and implicit reaction updates in a correction equation that when
applied iteratively achieves high-order time-accuracy.  In an operator
split implementation, reaction and burning proceed without knowing
about the other process.  In contrast, the update in the SDC
formulation explicitly couples the processes together.  In \castro\ we
consider reactions and hydrodynamics, and write our conservative
system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The conserved state has the form:
\begin{equation}
\Uc = (\rho, \rho X_k, \rho u, \rho v, \rho w, \rho E)^\intercal
\end{equation}
where $\rho$ is the mass density, $X_k$ are the mass fractions of the
nuclear species, $\Ub = (u, v, w)$ is the Cartesian velocity vector,
and $E$ is the total specific energy, related to the specific internal
energy, $e$, as $E = e + |\Ub|^2/2$.  The advective term takes the form:
\begin{equation}
\Advs{\Uc} = -\nabla \cdot \Fb (\Ub) + {\bf H}
\end{equation}
where $\Fb(\Ub)$ are the hydrodynamical fluxes:
\begin{equation}
\Fb \cdot \hat{\bf x} = \left (\begin{array}{c} \rho u \\ \rho X_k u \\ \rho u^2 + p \\ \rho v u \\ \rho w u \\ (\rho E + p) u  \end{array} \right )
\quad
\Fb \cdot \hat{\bf y} = \left (\begin{array}{c} \rho v \\ \rho X_k v \\ \rho u v \\ \rho v^2 + p \\ \rho w v \\ (\rho E + p) v  \end{array} \right )
\quad
\Fb \cdot \hat{\bf z} = \left (\begin{array}{c} \rho w \\ \rho X_k w \\ \rho u w \\ \rho v w \\ \rho w^2 + p \\ (\rho E + p) w  \end{array} \right )
\end{equation}
Here, $p$ is the pressure, which can be evaluated through our equation
of state, $p = p(\rho, e, X_k)$.  Note that we do not include
transport terms in this system, but it is straightforward to add in a
thermal diffusion flux.  ${\bf H}$ captures any hydrodynamical source
terms (e.g., gravity).  Finally the reactive sources take the form:
\begin{equation}
\Rbs{\Uc} = (0, \rho \omegadot_k, 0, 0, 0, \rho \dot{\epsilon})^\intercal
\end{equation}
where $\omegadot_k$ is the creation rate of species $k$ and $\dot{\epsilon}$
is the specific energy generation rate from reactions.

@@

\begin{table}[t]
\renewcommand{\arraystretch}{1.25}
\centering
\begin{tabular}{llllll}
{\bf field} & {$\epsilon_{64 \rightarrow 128}$} & 
{\bf rate} & {$\epsilon_{128\rightarrow 256}$} & 
{\bf rate} & {$\epsilon_{256\rightarrow 512}$} \\
\hline
\multicolumn{6}{c}{\bf Strang + CTU} \\
\hline
 $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.458 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.717 \times 10^{14}$  & 1.648  & $5.479 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
\hline
\multicolumn{6}{c}{\bf 2nd order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & $1.042 \times 10^{17}$  & 2.032  & $2.546 \times 10^{16}$  & 2.019  & $6.281 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.564 \times 10^{14}$  & 1.935  & $4.090 \times 10^{13}$  & 2.003  & $1.020 \times 10^{13}$  \\
\hline
\multicolumn{6}{c}{\bf 4th order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & $8.789 \times 10^{15}$  & 3.798  & $6.319 \times 10^{14}$  & 3.911  & $4.201 \times 10^{13}$  \\
 $\rho X(\isotm{O}{16})$     & $1.294 \times 10^{13}$  & 3.765  & $9.518 \times 10^{11}$  & 3.872  & $6.501 \times 10^{10}$  \\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1.0}
\caption{\label{table:convergence} Convergence of carbon mass density
  ($\rho X(\isotm{C}{12})$) and oxygen mass density ($\rho
  X(\isotm{O}{16})$) for the Strang-CTU algorithm and 2nd and 4th
  order SDC.  Data taken from \cite{castro:sdc}.}
\end{table}

>>>>

The SDC algorithm used in \castro\ follows the ideas of
\cite{dutt:2000,minion:2003}, and uses low order explicit advection
and implicit reaction updates in a correction equation that when
applied iteratively achieves high-order time-accuracy.  In an operator
split implementation, reaction and burning proceed without knowing
about the other process.  In contrast, the update in the SDC
formulation explicitly couples the processes together.  In \castro\ we
consider reactions and hydrodynamics, and write our conservative
system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The conserved state has the form:
\begin{equation}
\Uc = (\rho, \rho X_k, \rho u, \rho v, \rho w, \rho E)^\intercal
\end{equation}
where $\rho$ is the mass density, $X_k$ are the mass fractions of the
nuclear species, $\Ub = (u, v, w)$ is the Cartesian velocity vector,
and $E$ is the total specific energy, related to the specific internal
energy, $e$, as $E = e + |\Ub|^2/2$.  The advective term takes the form:
\begin{equation}
\Advs{\Uc} = -\nabla \cdot \Fb (\Uc) + {\bf H}
\end{equation}
where $\Fb(\Uc)$ are the hydrodynamical fluxes:
\begin{equation}
\Fb \cdot \hat{\bf x} = \left (\begin{array}{c} \rho u \\ \rho X_k u \\ \rho u^2 + p \\ \rho v u \\ \rho w u \\ (\rho E + p) u  \end{array} \right )
\quad
\Fb \cdot \hat{\bf y} = \left (\begin{array}{c} \rho v \\ \rho X_k v \\ \rho u v \\ \rho v^2 + p \\ \rho w v \\ (\rho E + p) v  \end{array} \right )
\quad
\Fb \cdot \hat{\bf z} = \left (\begin{array}{c} \rho w \\ \rho X_k w \\ \rho u w \\ \rho v w \\ \rho w^2 + p \\ (\rho E + p) w  \end{array} \right )
\end{equation}
Here, $p$ is the pressure, which can be evaluated through our equation
of state, $p = p(\rho, e, X_k)$.  Note that we do not include
transport terms in this system, but it is straightforward to add in a
thermal diffusion flux.  ${\bf H}$ captures any hydrodynamical source
terms (e.g., gravity).  Finally the reactive sources take the form:
\begin{equation}
\Rbs{\Uc} = (0, \rho \omegadot_k, 0, 0, 0, \rho \dot{\epsilon})^\intercal
\end{equation}
where $\omegadot_k$ is the creation rate of species $k$ and $\dot{\epsilon}$
is the specific energy generation rate from reactions.

@@

\begin{table}[t]
\renewcommand{\arraystretch}{1.25}
\centering
\begin{tabular}{lcc}
{\bf field} & {$\log_2 \frac{\epsilon_{64 \rightarrow 128}}{\epsilon_{128\rightarrow 256}}$} &
{$\log_2 \frac{\epsilon_{128\rightarrow 256}}{\epsilon_{256\rightarrow 512}}$} \\[0.5em]
\hline
\multicolumn{3}{c}{\bf Strang + CTU} \\
\hline
 $\rho X(\isotm{C}{12})$     & 1.945  & 2.194  \\
 $\rho X(\isotm{O}{16})$     & 1.648  & 1.898  \\
\hline
\multicolumn{3}{c}{\bf 2nd order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & 2.032  & 2.019  \\
 $\rho X(\isotm{O}{16})$     & 1.935  & 2.003  \\
\hline
\multicolumn{3}{c}{\bf 4th order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & 3.798  & 3.911  \\
 $\rho X(\isotm{O}{16})$     & 3.765  & 3.872  \\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1.0}
\caption{\label{table:convergence} Convergence of carbon mass density
  ($\rho X(\isotm{C}{12})$) and oxygen mass density ($\rho
  X(\isotm{O}{16})$) for the Strang-CTU algorithm and 2nd and 4th
  order SDC for a simple reacting acoustic pulse.  Here $\epsilon_{64\rightarrow 128}$ is the norm of the 
  difference between a $64^2$ and $128^2$ simulation, and similar for the other $\epsilon$s.  The convergence rate is the $\log_2$ of the ratio of errors between different resolutions.  Data taken from \cite{castro:sdc}.}
\end{table}
